---
title: "Trial match report"
author: "Printed from digital ECMT cancer trial matching tool"
date: ""
output: html_document
params: 
  cancer.type: NA
  include.solid: NA
  gene.list: NA
  drug.class.list: NA
  hide.cancer.alone: NA

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
require(flexdashboard)
require(dplyr)
require(tidyr)
require(kableExtra)
require(shinyWidgets)
require(leaflet)
require(htmltools)
require(rjson)


## load configuration data from JSON file
configuration <- fromJSON(file = "trialMatchConfiguration.json")
```

```{r get todays date}
##get today's date
today <- format(Sys.Date(), format = "%d %B %Y")

```
  
### **Search parameters**   
  
Report date: **`r today`**  

```{r get parameter values}
cancer.type <- params$cancer.type
include.solid <- params$include.solid
gene.list <- params$gene.list
drug.class.list <- params$drug.class.list[[1]]
hide.cancer.alone <- params$hide.cancer.alone

```

  
```{r read cancer study data into memory}
## read cancerStudies table into memory
# contains study details, mapped to a controlled set of cancer types in the "TARGET.condition"  column
# names are "locations, nct_id, brief_title, overall_status, condition, interventions, postcode, site_status, investigators, contacts, Refresh.date, matching.condition, TARGET.condition, Link, lat, long"
# no columns are aggregated
cancerStudies <- unique(read.table(file = "cancerStudies.tsv", header = TRUE, stringsAsFactors = FALSE))

## filter and retain only studies with overall status of "Recruiting"
cancerStudies <- dplyr::filter(cancerStudies, overall_status == "Recruiting")

## filter out sites with status of not yet recruiting
cancerStudies <- dplyr::filter(cancerStudies, site_status == "Recruiting")

## read conditions and synonyms into memory
conditionSynonyms <- read.csv(file="conditionSynonyms4.csv", header = TRUE, stringsAsFactors = FALSE)
```

```{r get a full set of all study locations}
cancer.study.locations <- unique(dplyr::select(cancerStudies, nct_id, lat, long))
## drop any rows with missing lat/long values
## NOTE THAT THIS MEANS THESE STUDIES WILL NOT BE DISPLAYED ON MAP
## (BUT WILL BE IN THE TABLE UNDERNEATH)
cancer.study.locations <- cancer.study.locations[complete.cases(cancer.study.locations), ]
```

```{r get mechanisms}
targetMatches <- unique(read.table(file = "targetMatches.tsv", header = TRUE, stringsAsFactors = FALSE))
## drop unneeded columns
targetMatches <- unique(dplyr::select(targetMatches, Interventions, Mechanism))

## join to cancerStudies table
cancerStudies <- merge(x=cancerStudies, by.x="interventions", all.x=TRUE, y=targetMatches, by.y="Interventions")
```


```{r aggregate cancerStudies into compact form}
## aggregating interventions was causing some study:intervention combinations to be missed out
cancerStudies <- cancerStudies %>%
       group_by_at(vars(-c(locations, postcode, condition,  site_status, investigators, contacts, lat, long, Refresh.date))) %>%
       summarize(locations = toString(sort(unique(na.omit(locations)))),
                 conditions = toString(sort(unique(na.omit(condition)))),
                 investigators = toString(sort(unique(na.omit(investigators)))),
                 contacts = toString(sort(unique(na.omit(contacts))))) %>%
       as.data.frame()
```


```{r read studies scored on match against genes of interest}
## scoredMatches contains all possible matches of studies, their interventions and inclusion criteria against all possible genes of interest
## column names are "nct_id, gene.of.interest, intervention.name, intervention.target, rationale, confidence, criteria"

scoredMatches <- unique(read.table(file = "scoredMatches.tsv", header = TRUE, stringsAsFactors = FALSE))

```
  
Selected cancer type: **`r cancer.type`**  
  
  
```{r get synonyms for selected cancer type}
synonyms <- paste(unique(conditionSynonyms$condition.synonyms[conditionSynonyms$controlled.cancer.type==cancer.type]), collapse = ", ")

solid.included <- ifelse(include.solid=="TRUE", yes = "", no = " not ")
```
  
The following synonyms for the selected cancer type were included when evaluating potential matches against clinical trials: **`r synonyms`**   
  
Matches against solid tumours **were `r solid.included` included** in this search.  
  
The following genes were used to rank studies: **`r gene.list`**  

```{r check whether selected mechanisms are present}
drug.class.label <- ifelse(!is.null(drug.class.list) & drug.class.list %in% cancerStudies$Mechanism, yes = drug.class.list, no = NA) 

```


Results were filtered to retain only studies that included the following mechanisms: **`r drug.class.label`**  
  
  
```{r check whether include matches on cancer type alone}

exclude.cancer.alone <- ifelse(hide.cancer.alone=="TRUE", yes = "excluded from", no = "included in")
```
    
**Matches based on cancer type alone have been `r exclude.cancer.alone` search results**.  
  
### **Search results**  
  
```{r generate table}

if(include.solid=="TRUE") {cancer.type <- c(cancer.type, "Solid")}

#cancer.type

cancerStudies <- dplyr::filter(cancerStudies, TARGET.condition %in% cancer.type)

if(!is.null(gene.list)) {
      scoredMatches <- dplyr::filter(scoredMatches, gene.of.interest %in% gene.list)
      cancerStudies <- unique(merge(x=cancerStudies, by.x=c("nct_id"), all.x=TRUE, 
                      y=scoredMatches, by.y=c("nct_id")))
      cancerStudies <- cancerStudies[order(cancerStudies$confidence, decreasing = TRUE), ]
} else cancerStudies$rationale <- NA
cancerStudies$rationale[is.na(cancerStudies$rationale)] <- "Match on cancer type alone"


if(!is.null(drug.class.list) & drug.class.list %in% cancerStudies$Mechanism) {
    cancerStudies <- dplyr::filter(cancerStudies, Mechanism %in% drug.class.list)
    cancerStudies$rationale <- paste0("Match on drug class (", cancerStudies$interventions, "; ", cancerStudies$Mechanism, ")")
}

if(hide.cancer.alone=="TRUE") {cancerStudies <- dplyr::filter(cancerStudies, rationale != "Match on cancer type alone")}


HTML(cancerStudies %>%
            mutate(nct_id = cell_spec(x=cancerStudies$nct_id,format="html", link = cancerStudies$Link)) %>%
            dplyr::select(-c(Link, TARGET.condition, conditions)) %>%
            mutate(locations = cell_spec(x=cancerStudies$locations,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           mutate(investigators = cell_spec(x=cancerStudies$investigators,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           mutate(contacts = cell_spec(x=cancerStudies$contacts,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           mutate(criteria = cell_spec(x=cancerStudies$criteria,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           dplyr::select("Study ID" = nct_id,
                         "Brief title" = "brief_title",
                         "Overall status" = "overall_status",
                         "Rationale" = "rationale",
                         "Matching cancer type" = "matching.condition",
                         "Matching inclusion criterion" = "criteria",
                         "Matching gene" = "gene.of.interest",
   #                      "Interventions" = "interventions",
   #                      Mechanism,
                         "Recruiting sites" = "locations",
                         "Investigators" = "investigators",
                         "Contacts" = "contacts"
                         ) %>%
            unique() %>%
            kable("html", escape = FALSE) %>%
            # kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
            kable_styling(bootstrap_options = c("hover", "condensed", "responsive")) %>%
            collapse_rows(columns = 1:3, valign = "top") %>%
            #row_spec(row=3, hline_after = TRUE) %>%
            kable_styling(fixed_thead = T))


```
  
    
  

