---
title: "digital ECMT cancer trial matching tool"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    orientation: rows
    logo: digitalECMTlogo48px.PNG
    theme: spacelab
    # includes:
    #    in_header: JS/matomo.js
runtime: shiny

---
```{r setup, include=FALSE}
require(flexdashboard)
require(dplyr)
require(tidyr)
require(kableExtra)
require(shinyWidgets)
require(leaflet)
require(htmltools)
require(rjson)


## load configuration data from JSON file
configuration <- fromJSON(file = "trialMatchConfiguration.json")
```
  
  
  
```{r read cancer study data into memory}
## read cancerStudies table into memory
# contains study details, mapped to a controlled set of cancer types in the "TARGET.condition"  column
# names are "locations, nct_id, brief_title, overall_status, condition, interventions, postcode, site_status, investigators, contacts, Refresh.date, matching.condition, TARGET.condition, Link, lat, long"
# no columns are aggregated
cancerStudies <- unique(read.table(file = "cancerStudies.tsv", header = TRUE, stringsAsFactors = FALSE))

## filter and retain only studies with overall status of "Recruiting"
cancerStudies <- dplyr::filter(cancerStudies, overall_status == "Recruiting")

## filter out sites with status of not yet recruiting
cancerStudies <- dplyr::filter(cancerStudies, site_status == "Recruiting")

## read conditions and synonyms into memory
conditionSynonyms <- read.csv(file="conditionSynonyms4.csv", header = TRUE, stringsAsFactors = FALSE)
```

```{r get a full set of all study locations}
cancer.study.locations <- unique(dplyr::select(cancerStudies, nct_id, lat, long))
## drop any rows with missing lat/long values
## NOTE THAT THIS MEANS THESE STUDIES WILL NOT BE DISPLAYED ON MAP
## (BUT WILL BE IN THE TABLE UNDERNEATH)
cancer.study.locations <- cancer.study.locations[complete.cases(cancer.study.locations), ]
```

```{r get mechanisms}
targetMatches <- unique(read.table(file = "targetMatches.tsv", header = TRUE, stringsAsFactors = FALSE))
## drop unneeded columns
targetMatches <- unique(dplyr::select(targetMatches, Interventions, Mechanism))

## join to cancerStudies table
cancerStudies <- merge(x=cancerStudies, by.x="interventions", all.x=TRUE, y=targetMatches, by.y="Interventions")
```


```{r aggregate cancerStudies into compact form}
## aggregating interventions was causing some study:intervention combinations to be missed out
cancerStudies <- cancerStudies %>%
       group_by_at(vars(-c(locations, postcode, condition,  site_status, investigators, contacts, lat, long, Refresh.date))) %>%
       summarize(locations = toString(sort(unique(na.omit(locations)))),
                 conditions = toString(sort(unique(na.omit(condition)))),
                 investigators = toString(sort(unique(na.omit(investigators)))),
                 contacts = toString(sort(unique(na.omit(contacts))))) %>%
       as.data.frame()
```


```{r read studies scored on match against genes of interest}
## scoredMatches contains all possible matches of studies, their interventions and inclusion criteria against all possible genes of interest
## column names are "nct_id, gene.of.interest, intervention.name, intervention.target, rationale, confidence, criteria"

scoredMatches <- unique(read.table(file = "scoredMatches.tsv", header = TRUE, stringsAsFactors = FALSE))


```

Inputs {.sidebar}
-----------------------------------------------------------------------

**1. Filter on cancer type**    

```{r  cancer type filter}
## display filter for cancer type
available.cancer.types <- sort(unique(cancerStudies$TARGET.condition))
selectInput(inputId = "cancerTypeFilter", label=NULL, choices = available.cancer.types, selected = NULL, multiple = FALSE,  selectize = TRUE, width = NULL, size = NULL)


```

<!-- **Cancer types included in search:**   -->

```{r reactive label}
synonymsLabel <- reactive({
   synonyms <- paste(unique(conditionSynonyms$condition.synonyms[conditionSynonyms$controlled.cancer.type==input$cancerTypeFilter]), collapse = ", ")
   synonyms
})

## uncomment this line if you want to show synonyms included in matches for each cancer type
# shiny::renderText({
#    paste(synonymsLabel(), collapse = "\n")
# })


```

  
```{r additional options}
checkboxInput(inputId="solidTumourFilter", label="Include matches on solid tumour?", value = TRUE, width = NULL)
# checkboxInput(inputId="notYetRecruitingFilter", label="Include sites not yet recruiting?", value = TRUE, width = NULL)
```

**2. Order results on genes of interest (optional)**    
 
```{r define static gene filter}
gene.list <- sort(unique(scoredMatches$gene.of.interest))
selectInput(inputId = "geneFilter", label=NULL, choices = gene.list, selected = NULL, multiple = TRUE,  selectize = TRUE, width = NULL, size = NULL)
```

**3. Filter results on drug classes (optional)**    


```{r define reactive mechanism filter}
## define how filter values will be populated
output$reactive.mechanism.filter <- renderUI({
   if(input$solidTumourFilter==FALSE) {cancer.types <- input$cancerTypeFilter}
   else {cancer.types <- c(input$cancerTypeFilter, "Solid")}
   
   table <- dplyr::filter(cancerStudies, TARGET.condition %in% cancer.types)   
   mechanism.list <- sort(unique(table$Mechanism))

   selectInput(inputId = "mechanismFilter", label=NULL, choices = mechanism.list, selected = NULL, multiple = TRUE,  selectize = TRUE, width = NULL, size = NULL)
})
# display filter
shiny::renderUI({
   uiOutput("reactive.mechanism.filter") 
})
```

  
```{r omit cancer type alone}
checkboxInput(inputId="confidenceFilter", label="Hide matches on cancer type alone?", value = FALSE, width = NULL)
```


```{r filter scoredMatches table}
reactiveScoredMatches <- reactive({
   table <- scoredMatches
   if(!is.null(input$geneFilter)) {
      table <- dplyr::filter(table, gene.of.interest %in% input$geneFilter)
   }
   ## return table
   unique(table)
})
```

```{r filter cancerStudies table}
reactiveCancerStudies <- reactive({
   if(input$solidTumourFilter==FALSE) {cancer.types <- input$cancerTypeFilter}
   else {cancer.types <- c(input$cancerTypeFilter, "Solid")}
   table <- dplyr::filter(cancerStudies, TARGET.condition %in% cancer.types)
   if(!is.null(input$geneFilter)) {
      table <- merge(x=table, by.x=c("nct_id"), all.x=TRUE, 
                     y=reactiveScoredMatches(), by.y=c("nct_id"))
      ## sort on confidence decreasing
      table <- table[order(table$confidence, decreasing = TRUE), ]
   } else {table[ , c("gene.of.interest","intervention.name","intervention.target","rationale","confidence","criteria")] <- NA}  ## add empty columns so can select and rename for display...
   if(!is.null(input$mechanismFilter)) {
      table <- dplyr::filter(table, Mechanism %in% input$mechanismFilter)
      table$rationale <- paste0("Match on drug class (", table$interventions, "; ", table$Mechanism, ")")
   }
   table$rationale[is.na(table$rationale)] <- "Match on cancer type alone"
   if(input$confidenceFilter==TRUE) {table <- dplyr::filter(table, rationale != "Match on cancer type alone")}
   
   ## return table
   unique(table)
})
```
  
   
  
Row {.tabset}
-----------------------------------------------------------------------

### Map view of sites

```{r print a map with study sites indicated}

## zoomed based on lat and long values in configuration file
## NOTE: layerIds must be unique (i.e. one study would only appear once)
## need to use combination of study ID and location to make layerIds unique
output$studiesmap <- renderLeaflet({
   studies <- unique(dplyr::select(reactiveCancerStudies(), nct_id, brief_title))
   locations <- merge(x=studies,by.x="nct_id", y=cancer.study.locations, by.y="nct_id")
   locations$uniqueId <- paste(locations$nct_id, locations$lat, locations$long, sep = "_")
   if(nrow(locations)>0) {
     leaflet() %>%
         addProviderTiles(providers$OpenStreetMap.Mapnik) %>%
         setView(lat = configuration$zoom.lat, lng = configuration$zoom.long, zoom = 6) %>%
         addMarkers(data=locations, 
                    lat = ~lat,
                 lng = ~long,
                 popup = ~brief_title,
                 layerId = ~uniqueId, 
                 clusterOptions = markerClusterOptions())
  } else {
     leaflet() %>%
         addProviderTiles(providers$OpenStreetMap.Mapnik) %>%
         setView(lat = configuration$zoom.lat, lng = configuration$zoom.long, zoom = 6)
  }
})

## render reactive map
shiny::renderUI({
   leafletOutput("studiesmap")
   
})

## create a reactive variable that will hold ID of a study selected on map
map.selection <- reactiveValues(id = NA)

## detect clicks on markers and use them to get study ID for marker that was clicked
observeEvent(input$studiesmap_marker_click, { 
   map.selection$id <- input$studiesmap_marker_click$id
   map.selection$id <- gsub(pattern = "_.*", replacement = "", x=map.selection$id)
  })

## detect click on map and use them to reset value of map.selection
observeEvent(input$studiesmap_click, {
   map.selection$id <- NA
  })
```
  
  

### Study details  

```{r render reactive cancerStudies table}
## render table
shiny::renderUI({
   table <- reactiveCancerStudies()
   table[is.na(table)] <- "-"
   if(!is.na(map.selection$id )) {
      table <- dplyr::filter(table, nct_id == map.selection$id )
   }
   if(nrow(table)>0) {
   HTML(table %>%
            mutate(nct_id = cell_spec(x=table$nct_id,format="html", link = table$Link)) %>%
            dplyr::select(-c(Link, TARGET.condition, conditions)) %>%
            mutate(locations = cell_spec(x=table$locations,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           mutate(investigators = cell_spec(x=table$investigators,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           mutate(contacts = cell_spec(x=table$contacts,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           mutate(criteria = cell_spec(x=table$criteria,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           dplyr::select("Study ID" = nct_id, 
                         "Brief title" = "brief_title", 
                         "Overall status" = "overall_status", 
                         "Rationale" = "rationale",
                         "Matching cancer type" = "matching.condition", 
                         "Matching inclusion criterion" = "criteria",
                         "Matching gene" = "gene.of.interest",
   #                      "Interventions" = "interventions", 
   #                      Mechanism, 
                         "Recruiting sites" = "locations", 
                         "Investigators" = "investigators", 
                         "Contacts" = "contacts"
                         ) %>% 
            unique() %>%
            kable("html", escape = FALSE) %>%
            # kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
            kable_styling(bootstrap_options = c("hover", "condensed", "responsive")) %>%
            collapse_rows(columns = 1:3, valign = "top") %>%
            #row_spec(row=3, hline_after = TRUE) %>%
            kable_styling(fixed_thead = T))
            
   } else print("Please click on map to display details of matching studies")
})
```

  
### Study ranking details  

  
All studies are matched on cancer type.  
  
If a gene of interest, matching studies are ordered as follows:  

1. Gene of interest is mentioned in inclusion criteria, and it is the molecular target of the study drug.  

2. Gene of interest is mentioned in inclusion criteria, and the study drug targets a downstream gene.   
  
  
3. Gene of interest is mentioned in inclusion criteria.  
  
  
4. Study drug targets gene of interest.  
  
  
5. Study drug targets a downstream gene.  
  
  
6. Study matched on cancer type alone.  
  
### Acknowledgements
  
`r configuration$about.us`  
   
   
`r configuration$privacy.statement`  

  
This tool is a prototype, intended for research use by Medical Professionals. Please note that it is not a comprehensive source of information about clinical trials, and is not for use in clinical decision making. If you have any questions, please contact `r configuration$contact.email` for more information.  
  
    
* Trial data are sourced from clinicaltrials.gov via the Clinical Trials Transformation Initiative: https://www.ctti-clinicaltrials.org/  
  
* Data regarding mechanisms of trial interventions are sourced from the NCI Thesaurus: https://ncithesaurus.nci.nih.gov/ncitbrowser/  
  
* Data regarding biological pathways are sourced from KEGG (https://www.kegg.jp/kegg/) using the R BioConductor package: https://www.bioconductor.org/  
  
  
  

  
  


```{r exit script }
knitr::knit_exit()
```
  