---
title: "digital ECMT cancer trial matching tool: results"
#output: html_document
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    orientation: rows
    #logo: digitalECMTlogo48px.PNG
    theme: spacelab
params: 
  subject_ID: "Not provided"
  cancer_type: "cancer type"
  include_solid: "T/F"
  synonyms: "condition synonyms"
  genes: "mutations, fusions, amplifications"
  mechanism_filter: "mechanism(s)"
  hide_cancer_only: "T/F"
  table: "json format"
  privacy_statement: ""
  contact_email: ""
  country: ""
  about_us: ""
---


```{r copyright notice}
 # 
 # This file is part of the cancer-trial-match distribution (https://github.com/digital-ECMT/cancer-trial-match).
 # Copyright (C) 2021 digital ECMT
 # 
 # This program is free software: you can redistribute it and/or modify  
 # it under the terms of the GNU General Public License as published by  
 # the Free Software Foundation, version 3 or later.
 #
 # This program is distributed in the hope that it will be useful, but 
 # WITHOUT ANY WARRANTY; without even the implied warranty of 
 # MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU 
 # General Public License for more details.
 #
 # You should have received a copy of the GNU General Public License 
 # along with this program. If not, see <http://www.gnu.org/licenses/>.
 #


```




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(jsonlite)       ## MIT
require(formattable)    ## MIT
require(kableExtra)     ## MIT

##get today's date
today <- format(Sys.Date(), format = "%d %B %Y")

## load configuration data from JSON file
# configuration <- jsonlite::fromJSON(txt = "trialMatchConfiguration.json")

```
  

  

```{r load parameters}
## view parameters
## table is passed as a parameter in JSON format
#params$table

## get table from parameters and convert to data frame
paramTable <- as.data.frame(fromJSON(params$table))

paramTable <- as.data.frame(jsonlite::fromJSON(params$table))


#formattable(paramTable)

```
  
Row {.tabset}
-----------------------------------------------------------------------

    
### Study details   
  
```{r display table}

if(nrow(paramTable)>0) {
  table <- paramTable
   table[is.na(table)] <- "-"

   if(nrow(table)>0) {
   table <- aggregate(table, by=table['nct_id'], function(x) {paste(unique(x), collapse = ";\n")})
   table <- table[order(table$combined_score, decreasing = TRUE), ]
   datatable(dplyr::select(table,
                           "Brief title" = "brief_title",
                            # combined_score,
                            "Matching alteration(s)" = "gene_variant_type",
                            "Matching cancer type(s)" = "matching.condition",
                            "Eligibility" = "eligibility_rationale",
                            "Drug rationale" = "intervention_rationale",
                            "Est. completion date" = "primary_completion_date",
                            "Last updated" = "last_update_posted_date",
                            "Interventions" = "interventions",
                            "Matching inclusion criteria" = "matching_criteria",
                            "Prior Therapy exclusions" = Exclusions,
                            "Locations" = "locations",
                            # "Centres" = "sites",
                            # "Investigators" = "investigators",
                            # "Central contacts" = "central_contacts",
                            "Contacts" = "contacts",
                            # "StudyID" = nct_id,
                            # "Overall status" = "overall_status",
                            
                            # Mechanism
                            ),
                  rownames=FALSE, escape = 4, 
                  options = list(ordering=F,
                                 paging = TRUE,
                                 scrollX = TRUE,
                                 scrollY = "70vh",
                                 #scrollCollapse = TRUE,
                                 autoWidth = TRUE,
                                 columnDefs = list(list(width = '500px', targets = list(4)))
                                )
  )
     
     
     
     
     
     
     
     
     
     # HTML(table %>%
     #      mutate(brief_title = cell_spec(x=table$brief_title,format="html", link = table$Link, new_tab = TRUE )) %>%
     #        dplyr::select(-c(Link, TARGET.condition, conditions)) %>%
     #        mutate(locations = cell_spec(x=table$locations,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #      mutate(sites = cell_spec(x=table$sites,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #       mutate(investigators = cell_spec(x=table$investigators,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #       mutate(contacts = cell_spec(x=table$contacts,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #       mutate(criteria = cell_spec(x=table$matching_criteria,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #       mutate(Exclusions = cell_spec(x=table$Exclusions,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #      aggregate( by=table['nct_id'], function(x) {paste(unique(x), collapse = ";\n")}) %>%
     #      dplyr::select(combined_score,
     #                    "Matching cancer type(s)" = "matching.condition",
     #                     "Matching alteration(s)" = "gene_variant_type",
     #                     "Drug rationale" = "intervention_rationale",
     #                     "Eligibility" = "eligibility_rationale",
     #                     "Brief title" = "brief_title",
     #                    "Interventions" = "interventions",
     #                     "Matching inclusion criteria" = "criteria",
     #                    "Prior Therapy exclusions" = Exclusions,
     #                     "Locations" = "locations",
     #                    # "Centres" = "sites",
     #                    # "Investigators" = "investigators",
     #                     "Contacts" = "contacts",
     #                    # "Study ID" = nct_id
     #                    # "Overall status" = "overall_status"
     #                    # Mechanism
     #                     ) %>%
     #        arrange( desc(combined_score)) %>% ## sort on score desc
     #        dplyr::select(-combined_score) %>%
     #        unique() %>%
     #        kable("html", escape = FALSE) %>%
     #        # kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
     #        kable_styling(bootstrap_options = c("hover", "condensed", "responsive")) %>%
     #        #collapse_rows(columns = c(1,2,4,5), valign = "middle") %>%
     #        #row_spec(row=3, hline_after = TRUE) %>%
     #        kable_styling(fixed_thead = T))

   }
} else print("No matching studies based on selected options")






```
  
### Search details  
  
Date of report: **`r today`**   
  
Cancer type(s): **`r params$cancer_type`**  
  
Synonyms included for selected cancer type(s): **`r params$synonyms`**  
  
Include matches on solid tumour? **`r params$include_solid`**  
    
Genetic alterations included in search: **`r paste(params$genes, collapse=", ")`**  
  
Drug mechanism filters: **`r paste(params$mechanism_filter, collapse=", ")`**  
  
Hide matches on cancer type alone: **`r params$hide_cancer_only`**  
  
### About  
  
#### About us  

`r params$about_us`.    


We would be grateful for feedback as to utility of this tool – if you have any feedback, or you would like to know more detail about this tool, please contact `r params$contact_email`.  

#### What does this tool do?  
  
The digital ECMT cancer trial matching tool is intended to support clinicians in the matching of people with cancer to clinical trials based on their cancer type, genetic alterations and/or trial drug mechanism. This is an early version that could possibly change.  
  
  
Potential matching trials are ranked according to the following order:   
  
  
1.	Study is enrolling patients with specified genetic alterations, AND specified gene is targeted by a study drug.  

2.	Study is enrolling patients with specified genetic alterations, AND a study drug targets a gene immediately downstream.  

3.	Study is enrolling patients with specified genetic alterations.  

4.	Study drug targets gene of interest.  

5.	Study drug targets a gene immediately downstream of specified genetic alteration.  

6.	Match on cancer type alone.  
  
  
#### Data sources  
  
Trial data are sourced exclusively from clinicaltrials.gov via the Clinical Trials Transformation Initiative (https://www.ctti-clinicaltrials.org/ ).   

Data regarding mechanisms of trial interventions are sourced from the NCI Thesaurus: https://ncithesaurus.nci.nih.gov/ncitbrowser/  

Data regarding biological pathways are sourced from KEGG (https://www.kegg.jp/kegg/) using the R BioConductor package: https://www.bioconductor.org/  
  
#### Limitations  
  
**Please note that it is not a comprehensive source of information about clinical trials, and its utility not been validated for use in clinical decision making.**  

Trials that are not included in clinicaltrials.gov will not be included in the search.  
  
  
Only studies with at least one site in `r params$country` are included. If you are interested in hosting a site for other countries, please contact `r params$contact_email`.  
  
Data are refreshed daily, but clinicians are advised to contact the relevant investigators to confirm recruitment status.  

Study eligibility criteria are tagged with the relevant genetic alterations based on natural language processing – some criteria may be incorrectly tagged, and some relevant criteria may be missed.  
  
  
### Legal  
  
  
    
* The University of Manchester digital ECMT team’s cancer trial matching tool includes content intended for use only by healthcare professionals. This tool does not give professional advice; physicians and other healthcare professionals who use this tool should exercise their own clinical judgment as to the information it provides. Individuals with any type of medical condition are specifically cautioned to seek professional medical advice before beginning any sort of health treatment. For medical concerns, including decisions about medications and other treatments, non-medical users should always consult their physician or other qualified healthcare professional.  


*	We do not give medical advice, nor do we provide medical or diagnostic services. Medical information changes rapidly. Neither we nor our content providers guarantee that the content of this tool covers all possible uses, directions, precautions, drug interactions, or adverse effects that may be associated with any therapeutic treatments.  

*	In the absence of any negligence or other breach of duty by us: (1) your reliance upon information and content obtained by you at or through this site is solely at your own risk; and (2) neither we nor our content providers are responsible for any damage or injury (including death) to you, other persons, or property arising from any use of any product, information, idea, or instruction in the content of the tool.  
  
#### Privacy  
  
`r params$privacy_statement`.  
 
  

      
  
