---
title: "digital ECMT cancer trial matching tool: results"
#output: html_document
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    orientation: rows
    #logo: digitalECMTlogo48px.PNG
    theme: spacelab
params: 
  subject_ID: "Not provided"
  cancer_type: "cancer type"
  include_solid: "T/F"
  synonyms: "condition synonyms"
  genes: "mutations, fusions, amplifications"
  mechanism_filter: "mechanism(s)"
  hide_cancer_only: "T/F"
  table: "json format"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(rjson)
require(formattable)
require(kableExtra)

##get today's date
today <- format(Sys.Date(), format = "%d %B %Y")
```
  

  

```{r load parameters}
## view parameters
## table is passed as a parameter in JSON format
#params$table

## get table from parameters and convert to data frame
paramTable <- as.data.frame(fromJSON(params$table))

#formattable(paramTable)

```
  
Row {.tabset}
-----------------------------------------------------------------------

    
### Study details   
  
```{r display table}

if(nrow(paramTable)>0) {
  table <- paramTable
   table[is.na(table)] <- "-"

   if(nrow(table)>0) {
   table <- aggregate(table, by=table['nct_id'], function(x) {paste(unique(x), collapse = ";\n")})
   table <- table[order(table$combined_score, decreasing = TRUE), ]
   datatable(dplyr::select(table,
                            # combined_score,
                            "Matching cancer type(s)" = "matching.condition",
                            "Matching alteration(s)" = "gene_variant_type",
                            "Drug rationale" = "intervention_rationale",
                            "Eligibility" = "eligibility_rationale",
                            "Brief title" = "brief_title",
                            "Interventions" = "interventions",
                            "Matching inclusion criteria" = "matching_criteria",
                            "Prior Therapy exclusions" = Exclusions,
                            "Locations" = "locations",
                            # "Centres" = "sites",
                            # "Investigators" = "investigators",
                            "Contacts" = "contacts"
                            # "StudyID" = nct_id,
                            # "Overall status" = "overall_status",
                            
                            # Mechanism
                            ),
                  rownames=FALSE, escape = 4, 
                  options = list(ordering=F,
                                 paging = TRUE,
                                 scrollX = TRUE,
                                 scrollY = "70vh",
                                 #scrollCollapse = TRUE,
                                 autoWidth = TRUE,
                                 columnDefs = list(list(width = '500px', targets = list(4)))
                                )
  )
     
     
     
     
     
     
     
     
     
     # HTML(table %>%
     #      mutate(brief_title = cell_spec(x=table$brief_title,format="html", link = table$Link, new_tab = TRUE )) %>%
     #        dplyr::select(-c(Link, TARGET.condition, conditions)) %>%
     #        mutate(locations = cell_spec(x=table$locations,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #      mutate(sites = cell_spec(x=table$sites,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #       mutate(investigators = cell_spec(x=table$investigators,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #       mutate(contacts = cell_spec(x=table$contacts,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #       mutate(criteria = cell_spec(x=table$matching_criteria,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #       mutate(Exclusions = cell_spec(x=table$Exclusions,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
     #      aggregate( by=table['nct_id'], function(x) {paste(unique(x), collapse = ";\n")}) %>%
     #      dplyr::select(combined_score,
     #                    "Matching cancer type(s)" = "matching.condition",
     #                     "Matching alteration(s)" = "gene_variant_type",
     #                     "Drug rationale" = "intervention_rationale",
     #                     "Eligibility" = "eligibility_rationale",
     #                     "Brief title" = "brief_title",
     #                    "Interventions" = "interventions",
     #                     "Matching inclusion criteria" = "criteria",
     #                    "Prior Therapy exclusions" = Exclusions,
     #                     "Locations" = "locations",
     #                    # "Centres" = "sites",
     #                    # "Investigators" = "investigators",
     #                     "Contacts" = "contacts",
     #                    # "Study ID" = nct_id
     #                    # "Overall status" = "overall_status"
     #                    # Mechanism
     #                     ) %>%
     #        arrange( desc(combined_score)) %>% ## sort on score desc
     #        dplyr::select(-combined_score) %>%
     #        unique() %>%
     #        kable("html", escape = FALSE) %>%
     #        # kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
     #        kable_styling(bootstrap_options = c("hover", "condensed", "responsive")) %>%
     #        #collapse_rows(columns = c(1,2,4,5), valign = "middle") %>%
     #        #row_spec(row=3, hline_after = TRUE) %>%
     #        kable_styling(fixed_thead = T))

   }
} else print("No matching studies based on selected options")






```
  
### Search details  
  
Date of report: **`r today`**   
  
Cancer type(s): **`r params$cancer_type`**  
  
Synonyms included for selected cancer type(s): **`r params$synonyms`**  
  
Include matches on solid tumour? **`r params$include_solid`**  
    
Genetic alterations included in search: **`r paste(params$genes, collapse=", ")`**  
  
Drug mechanism filters: **`r paste(params$mechanism_filter, collapse=", ")`**  
  
Hide matches on cancer type alone: **`r params$hide_cancer_only`**  
  
  

      
  
### Study ranking details  

  
* Studies are filtered on cancer type, plus/minus solid tumour.  
  
* If one or more drug mechanisms are selected, only studies that include one or more matching interventions are displayed.  
  
* If one or more genetic alterations are selected, only studies relevant for those alterations are displayed, matches are ranked based on:  

    * Study is enrolling patients with specified genetic alterations, and specified gene is targeted by a study drug.  

    * Study is enrolling patients with specified genetic alterations, and a study drug targets a gene immediately downstream.   
  
  
    * Study is enrolling patients with specified genetic alterations.   
  
  
    * Study drug targets gene of interest.  
  
  
    * Study drug targets a gene immediately downstream of specified genetic alteration.  
    
  
### Acknowledgements
  
INTENDED USE: This tool is a prototype, intended for research use by Medical Professionals. Please note that it is not a comprehensive source of information about clinical trials, and is not for use in clinical decision making. 
  
    
DATA SOURCES:  
  
  
* Trial data are sourced from clinicaltrials.gov via the Clinical Trials Transformation Initiative: https://www.ctti-clinicaltrials.org/  
  
* Data regarding mechanisms of trial interventions are sourced from the NCI Thesaurus: https://ncithesaurus.nci.nih.gov/ncitbrowser/  
  
* Data regarding biological pathways are sourced from KEGG (https://www.kegg.jp/kegg/) using the R BioConductor package: https://www.bioconductor.org/  

    