---
title: "digital ECMT cancer trial matching tool: results"
#output: html_document
output: 
  flexdashboard::flex_dashboard:
    #source_code: embed
    orientation: rows
    #logo: digitalECMTlogo48px.PNG
    theme: spacelab
params: 
  subject_ID: "Not provided"
  cancer_type: "cancer type"
  include_solid: "T/F"
  genes: "mutations, fusions, amplifications"
  mechanism_filter: "mechanism(s)"
  hide_cancer_only: "T/F"
  table: "json format"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(rjson)
require(formattable)
require(kableExtra)

##get today's date
today <- format(Sys.Date(), format = "%d %B %Y")
```
  

  

```{r load parameters}
## view parameters
## table is passed as a parameter in JSON format
#params$table

## get table from parameters and convert to data frame
paramTable <- as.data.frame(fromJSON(params$table))

#formattable(paramTable)

```
  
Row {.tabset}
-----------------------------------------------------------------------

    
### Study details   
  
```{r display table}
# if(nrow(paramTable)>0) {
#   table <- paramTable
#   table[is.na(table)] <- "-"
#   ## convert rationale column to factor to enable filtering with dropdown list 
#   #table$rationale <- as.factor(table$rationale)
#   ## do same for gene.of.interest
#   table$gene.of.interest <- as.factor(table$gene.of.interest)
#   table$nct_id <- paste0('<a href="',table$Link,'" target="_blank">',table$nct_id ,  " </a>")
#   #table <- table[, -c('Link')]
#   table <- unique(dplyr::select(table, "Matching cancer type" = "matching.condition",
#                          "Matching gene" = "gene.of.interest",
#                          "Rationale" = "rationale",
#                          "Brief title" = "brief_title",
#                          "Matching inclusion criterion" = "criteria",
#                          "Recruiting sites" = "locations",
#                          "Investigators" = "investigators",
#                          "Contacts" = "contacts",
#                          "Study ID" = nct_id
#                          #"Overall status" = "overall_status"
#    #                      "Interventions" = "interventions",
#    #                      Mechanism
#                          ))
#   DT::datatable(table, escape = which(names(table) == "Study ID"), filter = 'top',autoHideNavigation = TRUE,
#           # selection = list(mode = 'multiple', selected = NULL, target = 'column'),
#           options=list(ordering=FALSE, 
#                        autoWidth=TRUE,
#                        columnDefs = list(list(width='200px', targets = 4))))
#  } else print("No matching studies found")



if(nrow(paramTable)>0) {
  table <- paramTable
   table[is.na(table)] <- "-"
   
   if(nrow(table)>0) {
   HTML(table %>%
            mutate(brief_title = cell_spec(x=table$brief_title,format="html", link = table$Link, new_tab = TRUE )) %>%
            dplyr::select(-c(Link, TARGET.condition, conditions)) %>%
            mutate(locations = cell_spec(x=table$locations,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
          mutate(sites = cell_spec(x=table$sites,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           mutate(investigators = cell_spec(x=table$investigators,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           mutate(contacts = cell_spec(x=table$contacts,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
           mutate(criteria = cell_spec(x=table$matching_criteria,format="html", extra_css = "display: -webkit-box;-webkit-line-clamp: 5; -webkit-box-orient: vertical;  overflow: auto;")) %>%
          aggregate( by=table['nct_id'], function(x) {paste(unique(x), collapse = ";\n")}) %>%
          dplyr::select("Matching cancer type(s)" = "matching.condition",
                         "Matching alteration(s)" = "gene_variant_type",
                         "Drug rationale" = "intervention_rationale",
                         "Eligibility" = "eligibility_rationale",
                         "Brief title" = "brief_title",
                         "Matching inclusion criteria" = "criteria",
                         "Locations" = "locations",
                        "Centres" = "sites",
                         "Investigators" = "investigators",
                         "Contacts" = "contacts",
                        # "Study ID" = nct_id
                         #"Overall status" = "overall_status"
   #                      "Interventions" = "interventions",
   #                      Mechanism
                         ) %>%
            unique() %>%
            kable("html", escape = FALSE) %>%
            # kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
            kable_styling(bootstrap_options = c("hover", "condensed", "responsive")) %>%
            #collapse_rows(columns = c(1,2,4,5), valign = "middle") %>%
            #row_spec(row=3, hline_after = TRUE) %>%
            kable_styling(fixed_thead = T))

   } else print("No matching studies found")
}






```
  
### Search details  
  
Date of report: **`r today`**   
  
Subject ID: **`r params$subject_ID`**  

Cancer type: **`r params$cancer_type`**  
  
Include matches on solid tumour? **`r params$include_solid`**  
  
Genes included in search: **`r paste(params$genes, collapse=", ")`**  
  
Drug mechanism filters: **`r paste(params$mechanism_filter, collapse=", ")`**  
  

      
  
### Study ranking details  

  
* Studies are filtered on cancer type, plus/minus solid tumour.  
  
* If one or more drug mechanisms are selected, only studies that include one or more matching interventions are displayed.  
  
* If one or more genetic alterations are selected, only studies relevant for those alterations are displayed, matches are ranked based on:  

    * Study is enrolling patients with specified genetic alterations, and specified gene is targeted by a study drug.  

    * Study is enrolling patients with specified genetic alterations, and a study drug targets a gene immediately downstream.   
  
  
    * Study is enrolling patients with specified genetic alterations.   
  
  
    * Study drug targets gene of interest.  
  
  
    * Study drug targets a gene immediately downstream of specified genetic alteration.  
    
  
### Acknowledgements
  
INTENDED USE: This tool is a prototype, intended for research use by Medical Professionals. Please note that it is not a comprehensive source of information about clinical trials, and is not for use in clinical decision making. 
  
    
DATA SOURCES:  
  
  
* Trial data are sourced from clinicaltrials.gov via the Clinical Trials Transformation Initiative: https://www.ctti-clinicaltrials.org/  
  
* Data regarding mechanisms of trial interventions are sourced from the NCI Thesaurus: https://ncithesaurus.nci.nih.gov/ncitbrowser/  
  
* Data regarding biological pathways are sourced from KEGG (https://www.kegg.jp/kegg/) using the R BioConductor package: https://www.bioconductor.org/  
    